---
title: "First approach to models"
author: "Daniela Linero"
date: "2/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

```{r load libraries, include=FALSE, warning=FALSE}
library(yarg) # useful functions for dealing with PREDICTS data
library(roquefort) # useful PREDICTS functions, particularly for plotting
library(raster) # for dealing with spatial data
library(dplyr) # for handy data manipulation functions
library(tidyr) # ditto
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(stringr)
```

**1. Load the diversity metrics that have been calculated from a table that:**

* Had corrected abundance measures using sampling effort
* Had meged sites


```{r load table}
diversity3 <- readRDS("./output/intermediate_files/02_Analysis_walkthrough_Site_metrics.rds")
```

**2. I am going to merge the land-use classes "Primary forest" and "Primary non-forest". I'm also going to merge the use intensities "Intense" and "light". Finally, I will set the land-use or use intensity  as NA if they are:**

* Secondary vegetation (indeterminate age)
* Cannot decide 
* Urban 

```{r merge and drop}
diversity4 <- diversity3 %>%
  
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as
    # these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide get NA, urban too because it 
    # has only 40 sites
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Predominant_land_use = na_if(Predominant_land_use, "Urban"),

    
    # set reference levels
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
    Use_intensity = factor(Use_intensity),
    Use_intensity = relevel(Use_intensity, ref = "Minimal use")
  )
```



```{r diversity5}
diversity5 <- diversity4 %>%  
  mutate(Use_intensity = str_replace_all(Use_intensity,
                                         pattern = c("Intense use" = "Intense light use",
                                                     "Light use" = "Intense light use")), 
         Use_intensity = na_if(Use_intensity, "Cannot decide"),
         Use_intensity = factor(Use_intensity),
         Use_intensity = relevel(Use_intensity, ref = "Minimal use")) 

```


**3. Test for collinearity**

```{r collinearity}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

corvif(diversity5[ , c("Predominant_land_use", "Use_intensity")])
```

**4. Get complete cases, that means dropping the rows that have NA in the columns of total abundance, predominant land use and use intensity**

```{r complete cases model data}

model_data <- drop_na(diversity5, 
                      Total_abundance, Predominant_land_use, Use_intensity) %>% 
  droplevels()
```



```{r }
table(model_data$Predominant_land_use, model_data$Use_intensity)
```

**5. Transform abundance measures**

Abundance data usually display a nonnormal error distribution because they have a positive mean-variance relationship and are zero-inflated (Purvis et al., 2018).  Given that some abudance measures are not integers (some are relative abudance or densities), I am not going to model the abudance with a Poisson distribution, but I'm going to transform it in order to meet the assumptions of linear mixed models. 

```{r transform abundance}
model_data1 <- model_data %>% mutate(logAbundance = log(RescaledAbundance + 1),
                     sqrtAbundance = sqrt(RescaledAbundance)
)
```

**6. Data explorations**

```{r plots}
# Distribution of the log rescaled total abundance
hist(model_data1$logAbundance)
```


```{r plotss}
# Distribution of the square root of the rescaled total abundance
hist(model_data1$sqrtAbundance)
```


```{r plotsss}
# Boxplot showing rescaled total abundance differences between land-use types and intensities
ggplot(model_data1, aes(x=Predominant_land_use, y= RescaledAbundance, color= Use_intensity)) + 
  geom_boxplot() + theme(axis.text.x = element_text(size=9, angle=45))
```

Since the distribution looks a little bit more normal using square root, I am going to start with that as response variable

**7. Select random effects structure **

```{r model1}
First_model <- lmer(sqrtAbundance ~ Predominant_land_use + Use_intensity +
             Predominant_land_use:Use_intensity + 
             (1|SS) + (1|SSB), data = model_data1)

```

```{r model2}
Second_model <- lmer(sqrtAbundance ~ Predominant_land_use + Use_intensity +
             Predominant_land_use:Use_intensity + 
             (1+Predominant_land_use|SS) + (1|SSB), data = model_data1)
```

```{r model3}
Third_model <- lmer(sqrtAbundance ~ Predominant_land_use + Use_intensity + 
             Predominant_land_use:Use_intensity + 
             (1+Use_intensity|SS) + (1|SSB), data = model_data1)

```

Compare the model using the Akaike's Information Criterion 
```{r}
AIC(First_model, Second_model, Third_model)
```

I am going to select the last model as it didn't have a warning and has a lower AIC than the first one. 

**8. Select fixed effects structure**

```{r fixed effects}
Anova(Third_model)
```

Since the interaction is significant, I am going to leave the two explanatory variables.

**9. Plot residuals of the model**

```{r residuals 1}
plot(Third_model)
```

```{r residuals 2}
simulationOutput <- simulateResiduals(fittedModel = Third_model)
# Acces the qq plot
plotQQunif(simulationOutput)
# Plot the residuals against the predicted value 
plotResiduals(simulationOutput)

```

**10. Model estimates**

```{r summary1}

summary(Third_model)

```

According to the model estimates:

* The average square root of the rescaled abundance in Primary forest with minimal use is 0.536
* The average square root of the rescaled abundance in minimally-used Pasture is 0.14 higher. Meaning that the abundance in minimally-used Pasture is 0.54+0.14=0.68
* The average square root of the rescaled abundance in Intermediate secondary vegetation with an intense-light use is 0.09 lower. Meaning that the abundance in Intermediate secondary vegetation with an intense-light use is 0.54-0.09 = 0.45
* The average square root of the rescaled abundance in Plantation forest with an intense-light use is 0.13 lower. Meaning that the abundance in Plantation forest with an intense-light use is 0.54-0.13 = 0.41
* The average square root of the rescaled abundance in Crops with an intense-light use is 0.27 lower. Meaning that the abundance in Crops with an intense-light use is 0.54-0.13 = 0.27

**11. Plot the results**

```{r plot results 1}
roquefort::PlotErrBar(model = Third_model,
                      data = model_data5,
                      responseVar = "sqrt(Rescaled Abundance)",
                      seMultiplier = 1.96,
                      secdAge = TRUE,
                      logLink = "n",
                      catEffects = c("Predominant_land_use"),
                      forPaper = TRUE,
                      plotLabels = FALSE)

```


**12. Run the models with the log of Abundance**

```{r}
First_model2 <- lmer(logAbundance ~ Predominant_land_use + Use_intensity +
             Predominant_land_use:Use_intensity + 
             (1|SS) + (1|SSB), data = model_data1)
```

```{r}
Second_model2 <- lmer(logAbundance ~ Predominant_land_use + Use_intensity +
             Predominant_land_use:Use_intensity + 
             (1+Predominant_land_use|SS) + (1|SSB), data = model_data1)
```

```{r}
Third_model2 <- lmer(logAbundance ~ Predominant_land_use + Use_intensity + 
             Predominant_land_use:Use_intensity + 
             (1+Use_intensity|SS) + (1|SSB), data = model_data1)
```

```{r}
# Choose best random effects structure
AIC(First_model2, Second_model2, Third_model2)
```

```{r}
# See the significance of fixed variables
Anova(Third_model2)
```

```{r plot residuals 2}
plot(Third_model2)

# Simulate the residuals and plot them
simulationOutput1 <- simulateResiduals(fittedModel = Third_model2, plot = T)

# Acces the qq plot
plotQQunif(simulationOutput1)

# Plot the residuals against the predicted value 
plotResiduals(simulationOutput1)

```

```{r}
# model estimates
summary(Third_model2)
```

```{r plot results 2}
roquefort::PlotErrBar(model = Third_model2,
                      data = model_data5,
                      responseVar = "log Abundance",
                      seMultiplier = 1.96,
                      secdAge = TRUE,
                      logLink = "n",
                      catEffects = c("Predominant_land_use", "Use_intensity"),
                      forPaper = TRUE,
                      plotLabels = FALSE)

```


