---
title: "First approach to models"
author: "Daniela Linero"
date: "2/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(error = TRUE)
library(knitr)
library(kableExtra)
```

```{r load libraries, include=FALSE, warning=FALSE}
library(yarg) # useful functions for dealing with PREDICTS data
library(roquefort) # useful PREDICTS functions, particularly for plotting
library(raster) # for dealing with spatial data
library(dplyr) # for handy data manipulation functions
library(tidyr) # ditto
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(stringr)
library(kableExtra)
```

**1. Load the diversity metrics that have been calculated from a table that:**

* Had corrected abundance measures using sampling effort
* Had merged sites
* Does not have NaN


```{r load table}
diversity4 <- readRDS("./output/cleaned_data/02_Statistical_Analysis_Site_metrics.rds")
```

**2. I will fisrt explore the number of sites that have a known land use category**


```{r merge and drop}
second_model_data <- diversity4 %>%
  # make a level of Primary minimal
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation 
    # as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide are transformed into NA.
    # Urban too because it has only 40 sites. 
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Predominant_land_use = na_if(Predominant_land_use, "Urban"),
   
    
     # Give a shorter name to some land use categories 
    Predominant_land_use = str_replace_all(Predominant_land_use,
                                           pattern = c("Young_secondary vegetation" =
                                                         "YSV",
                                                       "Intermediate secondary vegetation" =
                                                         "YSV", 
                                                       "Mature secondary vegetation" = "MSV")),
    
    Predominant_land_use = factor(Predominant_land_use),
    Predominant_land_use = relevel(Predominant_land_use, ref = "Primary"),
  )


# Drop sites that don't have abundance measures or land-use data
second_model_data1 <- drop_na(second_model_data, 
                             Total_abundance, Predominant_land_use) %>% droplevels()


# Check the number of sites for animals and plants
addmargins(table(second_model_data1$Predominant_land_use,
                 second_model_data1$Use_intensity, second_model_data1$Kingdom), 2)


```


**3. Merge use intensities for those land use categories that don't have enough sites to separate them into different intensities**

* I will separate Primary into minimal, light and intense uses
* I will separate ISv into minimal and light-intense uses
* I will separate Plantation forest into minimal and light-intense uses
* Everything else I will merge all intensities


```{r merge some intensities}
LandUse_divide <- c("Primary", "ISV", "Plantation forest")

second_model_data2 <- second_model_data1 %>%
  
  mutate(
    
    # Drop the Cannot decide intensity levels for the land-use categories that have 
    # enough sites for the minimal, and light/intense
    Use_intensity = ifelse(Predominant_land_use %in% LandUse_divide & 
                             Use_intensity == "Cannot decide",
                           NA,
                           paste(Use_intensity)), 
    
    # Join the intensity levels of light and intense for Plantation forest and ISV 
    Use_intensity = ifelse((Predominant_land_use == "Plantation forest" |
                              Predominant_land_use == "ISV") & 
                             (Use_intensity == "Intense use" | 
                                Use_intensity == "Light use"),
      str_replace_all(Use_intensity, pattern = c("Intense use" = "Light-intense use",
                                                 "Light use" = "Light-intense use")),
      paste(Use_intensity)),
    
    # Merge all the intensity levels for those land-use categories that don't have 
    # enough sites in each land-use type/intensity combination
    Use_intensity = ifelse(Predominant_land_use %nin% LandUse_divide,
                           str_replace_all(Use_intensity, pattern = c("Intense use" =
                                                                        "All", "Light
                                                                      use" = "All",
                                                                      "Minimal use" 
                                                                      = "All", "Cannot
                                                                      decide" =
                                                                        "All")), 
                      
                           paste(Use_intensity)),
    

    # Paste the land-use classes and intensity levels
    LandUse = ifelse(Predominant_land_use != "NA" & Use_intensity != "NA",
                     paste(Predominant_land_use, Use_intensity),
                     NA),
    
    # set reference level
    LandUse = factor(LandUse),
    LandUse = relevel(LandUse, ref = "Primary Minimal use")
  )

```

**4. Test for collinearity**

Since I'm going to explore the collinearity between categorical variables, I'm going to use the Generalized variance Inflation Factors function provided by Zuur et al., (2009)


```{r collinearity}
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

corvif(second_model_data2[ , c("LandUse", "Kingdom")])
```

**4. Get complete cases, that means dropping the rows that have NA in the columns of total abundance and LandUse**

```{r complete cases model data}
# Create a table with complete cases
second_model_data2 <- drop_na(second_model_data2, 
                              Total_abundance, LandUse) %>% droplevels()

# Check number of sites
table(second_model_data2$LandUse, second_model_data2$Kingdom)
```


**5. Transform abundance measures**

Abundance data usually display a nonnormal error distribution because they have a positive mean-variance relationship and are zero-inflated (Purvis et al., 2018).  Given that some abudance measures are not integers (some are relative abudance or densities), I am not going to model the abudance with a Poisson distribution, but I'm going to transform it in order to meet the assumptions of linear mixed models. 

```{r transform abundance}
second_model_data2 <- mutate(second_model_data2, 
                            logAbundance = log(RescaledAbundance + 1),
                            sqrtAbundance = sqrt(RescaledAbundance)
)
```

**6. Select random effects structure **

To select the random-effects structure, we use the method recommended by (Zuur et al., 2009)  of taking the most complex fixed-effects structure, including all interactions, that will be tested in the second stage of modelling, and use it to compare different random-effects structures.


```{r model1}
m2_1 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
             (1|SS) + (1|SSB), data = second_model_data2)
```

```{r model2}
m2_2 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
             (1|SS) + (1|SSB) + (1|Source_ID), data = second_model_data2)
```

```{r model3}
m2_3 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
             (1+LandUse|SS) + (1|SSB), data = second_model_data2)
```


```{r model 5}
m2_5 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
             (1+Predominant_land_use|SS) + (1|SSB), data = second_model_data2)
```


Compare the models that converge using the Akaike's Information Criterion 
```{r Akaike 1}
AIC(m2_1, m2_2)
```


**8. Select fixed effects structure**

```{r fixed effects}
Anova(m2_2)
```

Since the interaction is significant, I am going to leave the two explanatory variables.

**9. Plot residuals of the model**

```{r residuals 1}
simulationOutput2 <- simulateResiduals(fittedModel = m2_2)
# Acces the qq plot
plotQQunif(simulationOutput2)
# Plot the residuals against the predicted value 
plotResiduals(simulationOutput2)
```


**10. Model estimates**

```{r summary1}

summary(m2_2)

```

Table with confidence intervals

```{r CI model 2}
CI_m2_2 <- read.csv("./output/summary_m2_2.csv", header = TRUE, sep = ";")
kable(CI_m2_2, format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")
```


**11. Plot the results**

```{r plot results 1}

#Read r code from a file which contains the function to make the plot
source("./R/PlotErrBar_interactions.R")
source("./R/PlotErrBar_interactions_modified.R")

# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = m2_2, resp = "Abundance", Effect1 = "LandUse",
                             Effect2 = "Kingdom",
                        ylims = c(-0.4,0.4), pointtype = c(16,17),blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), y = -0.38, labels = c("Primary minimal use", 
                                      "Cropland all uses", 
                                      "ISV light-intense use",
                                      "ISV minimal use",
                                      "MSV all uses",
                                      "Pasture all uses",
                                      "Plantation forest light-intense use",
                                      "Plantation forest minimal use", 
                                      "Primary intense use",
                                      "Primary light use",
                                      "YSV all uses"), srt = 18, cex= 0.7)

```


**12. Run the models with the log of Abundance**

```{r log first}
m2l_2 <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1|SS) + (1|SSB) + (1|Source_ID), data = second_model_data2)
```


```{r plot residuals 2}
simulationOutput2l <- simulateResiduals(fittedModel = m2l_2)
# Acces the qq plot
plotQQunif(simulationOutput2l)
# Plot the residuals against the predicted value 
plotResiduals(simulationOutput2l)

```

**13. Merge all land-use intensities**

```{r merge all land uses}

Third_model_data <- diversity4 %>%
  # make a level of Primary minimal
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation
    # as these aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide are transformed into NA.
    # Urban too because it has only 40 sites. 
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Predominant_land_use = na_if(Predominant_land_use, "Urban"),
    
    
    # Give a shorter name to some land use categories 
    Predominant_land_use = str_replace_all(Predominant_land_use, pattern = 
                                             c("Young secondary vegetation" = "YSV",
                                               
                                                                             "Intermediate secondary vegetation" = "ISV", 
                                               
                                                                             "Mature secondary vegetation" = "MSV")),
    
   
     # Merge all the intensity levels for all land-use categories 
    Use_intensity = ifelse(Use_intensity != "NA", "All", "NA"),
    
    # Paste the land-use classes and intensity levels
    LandUse = ifelse(Predominant_land_use != "NA" & Use_intensity != "NA",
                     paste(Predominant_land_use, Use_intensity),
                     NA),
    
    # set reference level
    LandUse = factor(LandUse),
    LandUse = relevel(LandUse, ref = "Primary All")
  )

```

**14. Test for collinearity**

```{r collinearity all merged intensities}

# Calculate the VIF
corvif(Third_model_data[ , c("LandUse", "Kingdom")])
```

**15. Complete cases**

```{r complete third model}
# Create a table with complete cases
Third_model_data2 <- drop_na(Third_model_data, 
                              Total_abundance, LandUse) %>% droplevels()

# Check number of sites
table(Third_model_data2$LandUse, Third_model_data2$Kingdom)

```

**16. Transform abundance measures**

```{r abundance third}

# Transform RescaledAbundance.
Third_model_data2 <- mutate(Third_model_data2, 
                             logAbundance = log(RescaledAbundance + 1),
                             sqrtAbundance = sqrt(RescaledAbundance)
)

```

**17. Select random effects structure **

```{r models random third 1}
m3_1 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1|SS) + (1|SSB), data = Third_model_data2)
```


```{r models random third 2}
m3_2 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1|SS) + (1|SSB) + (1|Source_ID), data = Third_model_data2)
```

```{r models random third 3}
m3_3 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1+LandUse|SS) + (1|SSB), data = Third_model_data2)
```

```{r models random third 4}
m3_5 <- lmer(sqrtAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1+Predominant_land_use|SS) + (1|SSB), data = Third_model_data2)
```

```{r AIC third}
# Compare the models that converged
AIC(m3_1, m3_2)
```

**18. Choose fixed effects structure**

```{r fixed third}

# See the significance of the terms in the model.
Anova(m3_2)

# The Land_use:kingdom interaction is significant, so I won’t remove this term 
# from the model. 

# compare the models that converged using all use intensities merged, and the models that converged
# separating the use-intensities for some land uses
AIC(m2_1, m2_2, m3_1, m3_2)
anova(m3_2, m2_1)
```

**19. Plot residuals**

```{r residuals third}
simulationOutput3 <- simulateResiduals(fittedModel = m3_2)
# Acces the qq plot
plotQQunif(simulationOutput3)
# Plot the residuals against the predicted value 
plotResiduals(simulationOutput3)
```


**20. Model estimates**

```{r estimates third}
summary(m3_2)
```


Table with confidence intervals

```{r CI model 3}
CI_m3_2 <- read.csv("./output/summary_m3_2.csv", header = TRUE, sep = ";")
kable(CI_m3_2, format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")

```



**21. Plot results**

```{r results third plot}
# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = m3_2, resp = "Abundance", Effect1 = "LandUse",
                             Effect2 = "Kingdom",
                             ylims = c(-0.4,0.4), pointtype = c(16,17),blackwhite = FALSE)

# Plot the x label
text(x = c(0.8, 1.8, 2.8, 3.8, 4.7, 5.7, 6.8), y = -0.38, labels = c("Primary", 
                                            "Cropland", 
                                            "ISV",                                          
                                            "MSV",
                                            "Pasture",
                                            "Plantation forest",
                                            "YSV all"), srt = 15, cex= 0.9)

```


