---
title: "02_Statistical_Analysis_Combination_Landuses"
author: "Daniela Linero"
date: "3/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(error = TRUE)
library(knitr)
library(kableExtra)
```

```{r load libraries, include=FALSE, warning=FALSE}
library(yarg) # useful functions for dealing with PREDICTS data
library(roquefort) # useful PREDICTS functions, particularly for plotting
library(dplyr) # for handy data manipulation functions
library(tidyr) # ditto
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(stringr) # to replace text
```

```{r}
# Read r code from a file which contains the function to make the plots
source("./R/PlotErrBar_interactions.R")
source("./R/PlotErrBar_interactions_modified.R")

```

In order to know which levels to use in all of the models:
I will the run the abundance, species richness and Simpson´s diversity models with the same land-use/intensities levels, add their AICs and compare which option is better.

```{r}
# Call the function that merges lan-uses and intensities
source("./R/02_Statistical_Analysis_merge_LandUses_Intensities.R")
```


```{r}
# Load richness data
diversity_all_richness <- readRDS("./output/cleaned_data/02_Statistical_Analysis_Richness_Site_metrics_animals_endooPlants_notendooPlants.rds")

# Load abundance data 
diversity_all_abundance <- readRDS(file = "./output/cleaned_data/02_Statistical_Analysis_Abundance_Site_metrics_combined_animals_endooPlants_notEndooPlants.rds")

# Transform rescaled abundance
diversity_all_abundance <- dplyr::mutate(diversity_all_abundance, 
                                  logAbundance = log(RescaledAbundance + 1)
)

# Load Simpsons data 
diversity_all_simpson <- readRDS("./output/cleaned_data/02_Statistical_Analysis_Simpson_Site_metrics_animals_endooPlants_notendooPlants.rds")

# Transform Simpson's diversity index
diversity_all_simpson <- diversity_all_simpson %>%
  
  # Create a new column to transform the response variable with log
  dplyr::mutate(log_one_over_D = log(Simpson_diversity))

```

*1. First option* 

* Primary: three level intensities 
*	Cropland:  minimal use and light/intense use 
* ISV: minimal use and light/intense use 
* MSV: merged
* Pasture: merged 
* Plantation forest: minimal use and light/intense 
* YSV: merged 

```{r}

# Create the vectors that hold the land-uses that we want to keep with 
# different use intensities 
land_uses_separate <- c("Primary", "Cropland", "ISV", "Plantation forest")
# Create a vector with the land-uses where we want to merge the light and intense use intensities
land_uses_light_intense <- c("Cropland", "ISV", "Plantation forest")


diversity_all_richness <- Merge_landUses_and_intensities(
  dataset = diversity_all_richness,
 index = 1, 
  land_uses_separate_intensities = land_uses_separate,
  land_uses_merge_light_intense = land_uses_light_intense,
  "Primary Minimal use")

diversity_all_abundance <- Merge_landUses_and_intensities(
  dataset = diversity_all_abundance,
  index = 1, 
  land_uses_separate_intensities = land_uses_separate,
  land_uses_merge_light_intense = land_uses_light_intense,
  "Primary Minimal use")

diversity_all_simpson <- Merge_landUses_and_intensities(
  dataset = diversity_all_simpson, 
  index = 1, 
  land_uses_separate_intensities = land_uses_separate,
   land_uses_merge_light_intense = land_uses_light_intense,
   "Primary Minimal use")
```

Species richness

```{r}

# Get complete cases, in order to compare models that use the same rows
diversity_richness <- drop_na(diversity_all_richness, 
                              Species_richness, LandUse.1)  %>%
  droplevels()

# Model richness with those land-use classes
m1_richness <- glmer(Species_richness ~ LandUse.1 + Kingdom + LandUse.1:Kingdom +
                (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS), 
                data = diversity_richness, 
                family = poisson, 
                control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 20000)))

# AIC
AIC(m1_richness)

```

Abundance

```{r}
# Get complete cases, in order to compare models that use the same rows
diversity_abundance <- drop_na(diversity_all_abundance, 
                               RescaledAbundance, LandUse.1) %>% droplevels()


# Model abundance with those land-use classes
m1_abundance <- lmer(logAbundance ~ LandUse.1 + Kingdom + LandUse.1:Kingdom +
                           (1|Source_ID) + (1|SS) + (1|SSB), 
                     data = diversity_abundance)

# AIC
AIC(m1_abundance)

```

Simpson's diversity index

```{r}
diversity_simpson <- drop_na(diversity_all_simpson, Simpson_diversity, LandUse.1)

# Model simpson's index with those land-use classes
m1_simpson <- lmer(log_one_over_D ~ LandUse.1 + Kingdom + LandUse.1:Kingdom +
                       (1|SS) + (1|SSB) + (1|Source_ID), data = diversity_simpson)

# AIC
AIC(m1_simpson)

```

*2. Second option* 

* Primary: three level intensities 
*	Cropland: merged
* ISV: minimal use and light/intense use 
* MSV: merged
* Pasture: merged 
* Plantation forest: minimal use and light/intense 
* YSV: merged 

```{r}
# Create the vectors that hold the land-uses that we want to keep with 
# different use intensities 
land_uses_separate_2 <- c("Primary", "ISV", "Plantation forest")
# Create a vector with the land-uses where we want to merge the light 
# and intense use intensities
land_uses_light_intense_2 <- c("ISV", "Plantation forest")


diversity_richness <- Merge_landUses_and_intensities(dataset = diversity_richness, 
                                                     index = 2, 
             land_uses_separate_intensities = land_uses_separate_2,
              land_uses_merge_light_intense = land_uses_light_intense_2,
               "Primary Minimal use")

diversity_abundance <- Merge_landUses_and_intensities(dataset = diversity_abundance,
                                                      index = 2, 
                     land_uses_separate_intensities = land_uses_separate_2,
                      land_uses_merge_light_intense = land_uses_light_intense_2,
                                                          "Primary Minimal use")

diversity_simpson <- Merge_landUses_and_intensities(dataset = diversity_simpson,
                                                    index = 2, 
                          land_uses_separate_intensities = land_uses_separate_2,
                          land_uses_merge_light_intense = land_uses_light_intense_2,
                                                     "Primary Minimal use")

```


Species richness

```{r}

# Model richness with those land-use classes
m2_richness <- glmer(Species_richness ~ LandUse.2 + Kingdom + LandUse.2:Kingdom +
                       (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS),
                     data = diversity_richness,
                     family = poisson, 
     control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 20000)))

# AIC
AIC(m2_richness)
```

Abundance

```{r}
# Model abundance with those land-use classes
m2_abundance <- lmer(logAbundance ~ LandUse.2 + Kingdom + LandUse.2:Kingdom +
                       (1|SS) + (1|SSB) + (1|Source_ID), data = diversity_abundance)

# AIC
AIC(m2_abundance)
```

Simpson's diversity index

```{r}

# Model simpson's index with those land-use classes
m2_simpson <- lmer(log_one_over_D ~ LandUse.2 + Kingdom + LandUse.2:Kingdom +
                     (1|SS) + (1|SSB) + (1|Source_ID), data = diversity_simpson)

# AIC
AIC(m2_simpson)
```


*3. Third option* 

* Primary:  minimal use and light/intense use  
*	Cropland: merged
* ISV: minimal use and light/intense use 
* MSV: merged
* Pasture: merged 
* Plantation forest: minimal use and light/intense 
* YSV: merged 

```{r}

# Create the vectors that hold the land-uses that we want to keep with different
# use intensities 
land_uses_separate_3 <- c("Primary", "ISV", "Plantation forest")
# Create a vector with the land-uses where we want to merge the light and 
# intense use intensities
land_uses_light_intense_3 <- c("Primary", "ISV", "Plantation forest")


diversity_richness <- Merge_landUses_and_intensities(dataset = diversity_richness, 
                                                     index = 3, 
                    land_uses_separate_intensities = land_uses_separate_3,
                    land_uses_merge_light_intense = land_uses_light_intense_3,
                                                     "Primary Minimal use")

diversity_abundance <- Merge_landUses_and_intensities(dataset = diversity_abundance, 
                                                      index = 3, 
                    land_uses_separate_intensities = land_uses_separate_3,
                    land_uses_merge_light_intense = land_uses_light_intense_3,
                                                      "Primary Minimal use")

diversity_simpson <- Merge_landUses_and_intensities(dataset = diversity_simpson,
                                                    index = 3, 
                    land_uses_separate_intensities = land_uses_separate_3,
                     land_uses_merge_light_intense = land_uses_light_intense_3,
                                                    "Primary Minimal use")

```

Species richness

```{r}
# Model richness with those land-use classes
m3_richness <- glmer(Species_richness ~ LandUse.3 + Kingdom + LandUse.3:Kingdom +
                       (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS), 
                     data = diversity_richness, 
                     family = poisson, 
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 20000)))


# AIC
AIC(m3_richness)

```

Abundance

```{r}
# Model abudance with those land-use classes
m3_abundance <- lmer(logAbundance ~ LandUse.3 + Kingdom + LandUse.3:Kingdom +
                       (1|SS) + (1|SSB) + (1|Source_ID),
                     data = diversity_abundance)

# AIC
AIC(m3_abundance)
```

Simpson's diversity

```{r}

# Model richness with those land-use classes
m3_simpson <- lmer(log_one_over_D ~ LandUse.3 + Kingdom + LandUse.3:Kingdom +
                     (1|SS) + (1|SSB) + (1|Source_ID),
                   data = diversity_simpson)

# AIC
AIC(m3_simpson)
```

*4. Fourth option* 

* Primary:  minimal use and light/intense use  
*	Cropland: minimal use and light/intense use
* ISV: minimal use and light/intense use 
* MSV: merged
* Pasture: merged 
* Plantation forest: minimal use and light/intense 
* YSV: merged 

```{r}
# Create the vectors that hold the land-uses that we want to keep with 
# different use intensities 
land_uses_separate_4 <- c("Primary","Cropland", "ISV", "Plantation forest")
# Create a vector with the land-uses where we want to merge the light 
# and intense use intensities
land_uses_light_intense_4 <- c("Primary", "Cropland", "ISV", "Plantation forest")


diversity_richness <- Merge_landUses_and_intensities(dataset = diversity_richness, 
                                                     index = 4, 
                    land_uses_separate_intensities = land_uses_separate_4,
                    land_uses_merge_light_intense = land_uses_light_intense_4,
                                                     "Primary Minimal use")

diversity_abundance <- Merge_landUses_and_intensities(dataset = diversity_abundance, 
                                                      index = 4, 
                    land_uses_separate_intensities = land_uses_separate_4,
                    land_uses_merge_light_intense = land_uses_light_intense_4,
                                                      "Primary Minimal use")

diversity_simpson <- Merge_landUses_and_intensities(dataset = diversity_simpson,
                                                    index = 4, 
                     land_uses_separate_intensities = land_uses_separate_4,
                    land_uses_merge_light_intense = land_uses_light_intense_4,
                                                    "Primary Minimal use")

```

Species richness

```{r}
# Model richness with those land-use classes
m4_richness <- glmer(Species_richness ~ LandUse.4 + Kingdom + LandUse.4:Kingdom +
                       (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS),
                     data = diversity_richness, family = poisson, 
   control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 20000)))


# AIC
AIC(m4_richness)
```

Abundance

```{r}
# Model abundance with those land-use classes
m4_abundance <- lmer(logAbundance ~ LandUse.4 + Kingdom + LandUse.4:Kingdom +
                       (1|SS) + (1|SSB) + (1|Source_ID), data = diversity_abundance)

# AIC
AIC(m4_abundance)
```

Simpson's diversity index

```{r}
# Model simpsons index with those land-use classes
m4_simpson <- lmer(log_one_over_D ~ LandUse.4 + Kingdom + LandUse.4:Kingdom +
                     (1|SS) + (1|SSB) + (1|Source_ID), data = diversity_simpson)

# AIC
AIC(m4_simpson)
```



*4. Choose the best combination*

```{r}
# First option
sum(AIC(m1_richness) + AIC(m1_abundance), AIC(m1_simpson))
```


```{r}
# Second option
sum(AIC(m2_richness) + AIC(m2_abundance), AIC(m2_simpson))
```


```{r}
# Third option
sum(AIC(m3_richness) + AIC(m3_abundance), AIC(m3_simpson))
```


```{r}
# Fourth option
sum(AIC(m4_richness) + AIC(m4_abundance), AIC(m4_simpson))

```

Acoording to the results, the best option is option 4:
* Primary:  minimal use and light/intense use  
*	Cropland: minimal use and light/intense use
* ISV: minimal use and light/intense use 
* MSV: merged
* Pasture: merged 
* Plantation forest: minimal use and light/intense 
* YSV: merged 

### The results therefore will be

#### Species richness

The best model is:
m4_richness <- glmer(Species_richness ~ LandUse.4 + Kingdom + LandUse.4:Kingdom +
                       (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS),
                     data = diversity_richness, family = poisson, 
   control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 20000)))


**Compare with a model that has all the land-use intensities merged**

```{r}
# Create the vectors that hold the land-uses that we want to keep with
# different use intensities 
land_uses_separate_null_5 <- "NA"
# Create a vector with the land-uses that we want to merge
land_uses_light_intense_null_5 <- "NA"


# This function creates new columns with the land-uses and use intensities that we want,
# then it then merges both columns into a LanUse column
diversity_richness <- Merge_landUses_and_intensities(diversity_richness, 
                                                5,
                                                land_uses_separate_null_5, 
                                                land_uses_light_intense_null_5,
                                                "Primary All")

# Model richness with those land-use classes
m4_null_1_richness <- glmer(Species_richness ~ LandUse.5 + Kingdom + LandUse.5:Kingdom +
                  (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS), 
                  data = diversity_richness,
                  family = poisson, 
        control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 20000)))


# Next we will check if we've lost a significant amount of explanatory power 
# by removing this interaction. If we have, we want to keep the more complex
# model. If we haven't lost a significant amount of explanatory power, then we
# can keep the simpler model. 

anova(m4_richness, m4_null_1_richness)


```

**Plant constrast**

```{r}
# For the null model I am going to replace the nePlantae for Plantae 
diversity_richness  <- diversity_richness  %>%

  mutate(
    
    # Create a new kingdom column as the copy of the kingdom column we used
    # in the first model
    Kingdom.1 = paste(Kingdom),
    
    # Replace nePlantae for Plantae
    Kingdom.1 = recode_factor(Kingdom.1, "nePlantae" = "Plantae"), 
    
    # set reference level
    Kingdom.1 = factor(Kingdom.1),
    Kingdom.1 = relevel(Kingdom.1, ref = "Animalia"))


m4_null_2_richness <- glmer(Species_richness ~ LandUse.4 + Kingdom.1 + LandUse.4:Kingdom.1 +
                  (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS),
                  data = diversity_richness, 
                  family = poisson, 
        control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 20000)))


# check if we've lost a significant amount of explanatory power 
anova(m4_richness, m4_null_2_richness, test = "F")

```

**Estimates of the selected model**

```{r }
estimates_richness <- read.csv("./output/summary_final_richness.csv", header = TRUE, sep = ";")
kable(estimates_richness, format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")
```

**Plot of the results**

```{r}

# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = m4_richness,
                             resp = "Species Richness",
                             Effect1 = "LandUse.4", 
                             Effect2 = "Kingdom",
                             ylims = c(-3,2.5),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -3, labels = c("Primary Min",
                                         "Cropland Li-In",
                                         "Cropland Min",
                                         "ISV Li-In",
                                         "ISV Min",
                                         "MSV All",
                                         "Pasture All",
                                         "Plantation Li-In",
                                         "Plantation Min",
                                         "Primary Li-In",
                                         "YSV All"),
     srt = 18, cex= 0.7)



```

#### Abundance

m4_abundance <- lmer(logAbundance ~ LandUse.4 + Kingdom + LandUse.4:Kingdom +
                       (1|SS) + (1|SSB) + (1|Source_ID), data = diversity_abundance)


**Compare with a model that has all the land-use intensities merged**

```{r}

# This function creates new columns with the land-uses and use intensities that we want,
# then it then merges both columns into a LanUse column
diversity_abundance <- Merge_landUses_and_intensities(diversity_abundance, 
                                                5,
                                                land_uses_separate_null_5, 
                                                land_uses_light_intense_null_5,
                                                "Primary All")

# Model abundance with those land-use classes
m4_null_1_abundance <- lmer(logAbundance ~ LandUse.5 + Kingdom +
                              LandUse.5:Kingdom +
                        (1|Source_ID) + (1|SS) + (1|SSB),
                        data = diversity_abundance)

# Next we will check if we've lost a significant amount of explanatory power 

anova(m4_abundance, m4_null_1_abundance)

```

**Plant constrast**

```{r}
# For the null model I am going to replace the nePlantae for Plantae 
diversity_abundance  <- diversity_abundance  %>%

  mutate(
    
    # Create a new kingdom column as the copy of the kingdom column we used
    # in the first model
    Kingdom.1 = paste(Kingdom),
    
    # Replace nePlantae for Plantae
    Kingdom.1 = recode_factor(Kingdom.1, "nePlantae" = "Plantae"), 
    
    # set reference level
    Kingdom.1 = factor(Kingdom.1),
    Kingdom.1 = relevel(Kingdom.1, ref = "Animalia"))


m4_null_2_abundance <- lmer(logAbundance ~ LandUse.4 + Kingdom.1 +
                              LandUse.4:Kingdom.1 +
                        (1|Source_ID) + (1|SS) + (1|SSB), 
                        data = diversity_abundance)


# check if we've lost a significant amount of explanatory power 
anova(m4_abundance, m4_null_2_abundance, test = "F")

```

**Estimates of the selected model**

```{r }
estimates_abundance <- read.csv("./output/summary_final_abundance.csv", header = TRUE, sep = ";")
kable(estimates_abundance, format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")
```

**Plot of the results**

```{r}

# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = m4_abundance,
                             resp = "Rescaled total abundance",
                             Effect1 = "LandUse.4", 
                             Effect2 = "Kingdom",
                             ylims = c(-0.2, 0.3),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -0.18, labels = c("Primary Min",
                                         "Cropland Li-In",
                                         "Cropland Min",
                                         "ISV Li-In",
                                         "ISV Min",
                                         "MSV All",
                                         "Pasture All",
                                         "Plantation Li-In",
                                         "Plantation Min",
                                         "Primary Li-In",
                                         "YSV All"),
     srt = 18, cex= 0.7)



```

### Simpson's diversity

m4_simpson <- lmer(log_one_over_D ~ LandUse.4 + Kingdom + LandUse.4:Kingdom +
                     (1|SS) + (1|SSB) + (1|Source_ID), data = diversity_simpson)


**Compare with a model that has all the land-use intensities merged**

```{r}

# This function creates new columns with the land-uses and use intensities that we want,
# then it then merges both columns into a LanUse column
diversity_simpson <- Merge_landUses_and_intensities(diversity_simpson, 
                                                5,
                                                land_uses_separate_null_5, 
                                                land_uses_light_intense_null_5,
                                                "Primary All")

# Model 1/D with those land-use classes
m4_null_1_simpson <- lmer(log_one_over_D ~ LandUse.5 + Kingdom + 
                            LandUse.5:Kingdom +
                        (1|Source_ID) + (1|SS) + (1|SSB), 
                        data = diversity_simpson)

# Next we will check if we've lost a significant amount of explanatory power 

anova(m4_simpson, m4_null_1_simpson)

```

**Plant constrast**

```{r}
# For the null model I am going to replace the nePlantae for Plantae 
diversity_simpson  <- diversity_simpson  %>%

  mutate(
    
    # Create a new kingdom column as the copy of the kingdom column we used
    # in the first model
    Kingdom.1 = paste(Kingdom),
    
    # Replace nePlantae for Plantae
    Kingdom.1 = recode_factor(Kingdom.1, "nePlantae" = "Plantae"), 
    
    # set reference level
    Kingdom.1 = factor(Kingdom.1),
    Kingdom.1 = relevel(Kingdom.1, ref = "Animalia"))


m4_null_2_simpson <- lmer(log_one_over_D ~ LandUse.4 + Kingdom.1 +
                            LandUse.4:Kingdom.1 +
                        (1|Source_ID) + (1|SS) + (1|SSB), 
                        data = diversity_simpson)


# check if we've lost a significant amount of explanatory power 
anova(m4_simpson, m4_null_2_simpson, test = "F")

```

**Estimates of the selected model**

```{r }
estimates_simpson <- read.csv("./output/summary_final_simpson.csv", header = TRUE, sep = ";")
kable(estimates_simpson, format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")
```

**Plot the results**

```{r}

# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = m4_simpson,
                             resp = "Simpson's diversity index",
                             Effect1 = "LandUse.4", 
                             Effect2 = "Kingdom",
                             ylims = c(-1, 1),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -0.9, labels = c("Primary Min",
                                         "Cropland Li-In",
                                         "Cropland Min",
                                         "ISV Li-In",
                                         "ISV Minimal",
                                         "MSV All",
                                         "Pasture All",
                                         "Plantation Li-In",
                                         "Plantation Min",
                                         "Primary Li-In",
                                         "YSV All"),
     srt = 18, cex= 0.7)

```
