---
title: "Statistical_Analysis_Simpson"
author: "Daniela Linero"
date: "1/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(error = TRUE)
library(knitr)
library(kableExtra)
```


```{r load libraries, include=FALSE, warning=FALSE}
library(yarg) # useful functions for dealing with PREDICTS data
library(roquefort) # useful PREDICTS functions, particularly for plotting
library(dplyr) # for handy data manipulation functions
library(tidyr) # ditto
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(stringr) # to replace text
```

*1. Load data*

Load the table that contains the abudance and occurrence measures for frugivores along with the plants dispersed and not dispersed by animals. 

```{r}
# Import table
diversityS <- readRDS(file = "./output/cleaned_data/01_Filter_data_frugi_endooPlants_notEndooPlants_records.rds")

```

*2. Correct abundance measures using Sampling effort*

```{r}
diversityS <- yarg::CorrectSamplingEffort(diversityS) 
```

*3. Select studies that assesed more than one species*

```{r}
# Create a dataset with the SS that assessed more than 1 species
list <- diversityS  %>% 
  
  # By grouping by SS we are comparing sites with the same diversity metric
  # type either abundance or occurence 
  dplyr::group_by(SS) %>%
  
  # Create a new column to calculate the number of species sampled per study 
  # n_distinct = length(unique())
  dplyr::mutate(N_species_sampled = dplyr::n_distinct(Taxon_name_entered)) %>%
  
  # ungroup data frame  
  dplyr::ungroup() %>%
  
  # Filter the studies that sampled more than 1 species
  dplyr::filter(N_species_sampled > 1) %>%
  
  base::droplevels()
```


```{r}
# Create a character vector of unique SS that assessed more than 1 species
list <- as.character(unique(list$SS))

diversityS <- diversityS %>%
  
  # Filter the studies present in the list
  base::subset(SS %in% list) %>% 
  
  droplevels()
```

*4. Remove sites that will produce NaN values*

```{r}
diversityS <- diversityS %>% 
  
  # Remove the sites that have a maximum abundance of 0
  base::subset(SSBS %nin% c("MJ1_2009__Lehouck 2 Fururu 10", 
                             "MJ1_2009__Lehouck 2 Fururu 11", 
                               "MJ1_2009__Lehouck 2 Macha 12", 
                                 "MJ1_2009__Lehouck 2 Macha 13")) %>%
  base::droplevels()

```

*5. Split dataset*

I will separate the measures of plants not dispersed by animals, in order to avoid
that they are merged together with plants dispersed by animals with the MergeSites function

```{r}
# I am going to separate the records that belong to plants not dispersed
# by animals
diversityS_notEndo <- diversityS %>%
  
  base::subset(Kingdom == "nePlantae") %>%
  
  base::droplevels()

# Get the table without plants dispersed by animals
diversityS_frugi_endo <- diversityS %>%
  
  base::subset(Kingdom != "nePlantae") %>%
  
  base::droplevels()

```

*6. Merge Sites*

```{r}
diversityS_frugi_endo <- yarg::MergeSites(diversityS_frugi_endo, 
                                          silent = TRUE, 
                                          merge.extra = "Wilderness_area")

diversityS_notEndo <- yarg::MergeSites(diversityS_notEndo, 
                                       silent = TRUE, 
                                       merge.extra = "Wilderness_area")

```

*7. Rename Predominant Habitat*

```{r}
# Rename the column predominant habitat, as the dataset is actually 
# refering to land use
diversityS_frugi_endo <-  dplyr::rename(diversityS_frugi_endo,
                        Predominant_land_use = Predominant_habitat)

diversityS_notEndo <-  dplyr::rename(diversityS_notEndo,
                    Predominant_land_use = Predominant_habitat)
```

*8. Calculate Diversity Metrics*
```{r, eval=FALSE}
# Calculate diversity metrics for animals and endoozoocoric plants
diversity1S_frugi_endo <- diversityS_frugi_endo %>%
  
  # add Diversity_metric_is_valid column
  dplyr::mutate(Diversity_metric_is_valid = TRUE) %>%
  
  # The extra.cols parameter is used for columns that we want to 
  # transferred to the final site-level data frame and that the function 
  # does not add  automatically
  yarg::SiteMetrics(extra.cols = c("SSB", "SSBS", "Predominant_land_use", "Kingdom")) 

# Calculate diversity metrics for plants not dispersed by animals
diversity1S_notendo <- diversityS_notEndo %>%
  
  # add Diversity_metric_is_valid column
  dplyr::mutate(Diversity_metric_is_valid = TRUE) %>%
  
  # The extra.cols parameter is used for columns that we want to 
  # transferred to the final site-level data frame and that the function 
  # does not add  automatically
  yarg::SiteMetrics(extra.cols = c("SSB", "SSBS", "Predominant_land_use", "Kingdom")) 

# Merge the site metrics for all organisms
diversity_all <- base::rbind.data.frame(diversity1S_frugi_endo,
                                        diversity1S_notendo)

```

```{r, include=FALSE}
# Import table
diversity_all <- readRDS("./output/cleaned_data/02_Statistical_Analysis_Simpson_Site_metrics_animals_endooPlants_notendooPlants.rds")

```

*9. Check results*

```{r}
# Merge the measures of frugivores, endoplants and not endoplants
diversityS_combined <- rbind.data.frame(diversityS_frugi_endo, 
                                        diversityS_notEndo)

diversityS_combined  %>% 
  
  # subset one site to check
  base::subset(SSS == "DG1_2013__Zou 1 1") %>% base::droplevels() %>%
  
  # Calculate each species abundance over total abundance at the site
  dplyr::mutate(Proportion = Measurement/sum(Measurement)) %>%
  
  # Sum the proportions and squared them
  dplyr::mutate(Simpson_D = sum(Proportion^2)) %>%
  
  # Calculate 1/D
  dplyr::mutate(one_over_D = 1/Simpson_D) %>% 
  
   # Select only columns we're interested in
  dplyr::select(SSS, Best_guess_binomial, Measurement, one_over_D)
```


```{r}
# Compare with the SiteMetrics result
diversity_all %>% base::subset(SSS == "DG1_2013__Zou 1 1") %>%
  dplyr::select(SSS, Simpson_diversity)


```

*10. Check the abundance measures units*

I am going to check that the measures used that weren’t number of individuals or ind/km2, were reasonably concordant to an abundance measure of each species. 


```{r}
# Get the unique diversity metric type and unit for each study
Diversity_metric_unit <- diversityS_combined %>% 

  # Get only one row for each study
  dplyr::distinct(SS, .keep_all = TRUE) %>%
  
  # Select the columns we're interested in
  dplyr::select(Source_ID, SS,  
                Diversity_metric_unit, Kingdom, 
                Study_common_taxon) %>%

  # Filter the studies that need to be checked
  base::subset(Diversity_metric_unit %nin% c("individuals",
                                             "stems/hectare", 
                                             "presence/absence", 
                                             "effort-corrected individuals",
                                             "individuals/km2", 
                                             "groups/colonies per km", 
                                             "number of groups",
                                             "individuals/km")) %>%
  droplevels()


kable(Diversity_metric_unit, format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")

```

*11. Models*

Count number of sites for the SiteMetrics table in order to know the sample size for each land-use type and intensity.


```{r}
# Remove sites that don't have a simpson diversity value or land-use type
diversity_simpson <- drop_na(diversity_all, 
                             Simpson_diversity, 
                             Predominant_land_use) %>%
  droplevels()

# Number of sites
table(diversity_simpson$Predominant_land_use,
                 diversity_simpson$Use_intensity,
                 diversity_simpson$Kingdom)
```


According to the number of sites, I am going to try this initial combination:

* Primary can be divided into the two level intensities in all cases
*	Cropland has to be merged
* ISV can be divided in minimal use and light/intense use 
* MSV has to be merged
* Pasture has to be merged 
* Plantation forest can be divided in minimal use and light/intense 
*	YSV has to be merged 


```{r}

# Call the function that merges lan-uses and intensities
source("./R/02_Statistical_Analysis_merge_LandUses_Intensities.R")

# Create the vectors that hold the land-uses that we want to
# keep with different use intensities 
land_uses_separate_1 <- c("Primary", "ISV", "Plantation forest")

# Create a vector with the land-uses where we want to merge the 
#light and intense use intensities
land_uses_light_intense_1 <- c("Primary", "ISV", "Plantation forest")

# Merge landuse intensities
diversity_simpson <- Merge_landUses_and_intensities(dataset = diversity_simpson,
                                                    index = 1, 
                       land_uses_separate_intensities = land_uses_separate_1,
                       land_uses_merge_light_intense = land_uses_light_intense_1,
                                                    "Primary Minimal use")

# Check number of sites
addmargins(table(diversity_simpson$LandUse.1, diversity_simpson$Kingdom), 2)
```

Test for collinearity

```{r}

# Get the function
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

# Calculate the VIF
corvif(diversity_simpson[ , c("LandUse.1", "Kingdom")])

```

Choose random effects structure

```{r}
# Simplest random effects structure 
m1 <- lmer(Simpson_diversity ~ LandUse.1 + Kingdom + LandUse.1:Kingdom +
             (1|SS) + (1|SSB), data = diversity_simpson) 
```


```{r}
plot(m1)
hist(diversity_simpson$Simpson_diversity)
# The simplest model does not converge

```

*12. Second attempt*

```{r}
# I am going to merge all categories

# Create the vectors that hold the land-uses that we want to keep
#with different use intensities 
land_uses_separate_2 <- "NA"
# Create a vector with the land-uses where we want to merge the
# light and intense use intensities
land_uses_light_intense_2 <- "NA"

diversity_simpson <- Merge_landUses_and_intensities(dataset = diversity_simpson,
                                                    index = 2, 
               land_uses_separate_intensities = land_uses_separate_2,
               land_uses_merge_light_intense = land_uses_light_intense_2,
                                                    "Primary All")

addmargins(table(diversity_simpson$LandUse.2, diversity_simpson$Kingdom), 2)
```

```{r}
m2 <- lmer(Simpson_diversity ~ LandUse.2 + Kingdom + LandUse.2:Kingdom +
             (1|SS) + (1|SSB), data = diversity_simpson) 
```

```{r}
plot(m2)
```

I am going to try to remove the MSV, since it only has 20 sites in total for nePlants

```{r}

# Remove MSV records
diversity_simpson1 <- diversity_simpson %>% 
  subset(LandUse.2 != "MSV All") %>% droplevels()

# Run the model
m3 <- lmer(Simpson_diversity ~ LandUse.2 + Kingdom + LandUse.2:Kingdom +
             (1|SS) + (1|SSB), data = diversity_simpson1) #Is Singular

```
```{r}
plot(m3)
```

