---
title: "Abundance model"
author: "Daniela Linero"
date: "17/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(error = TRUE)
library(knitr)
library(kableExtra)
```


```{r load libraries, include=FALSE, warning=FALSE}
library(yarg) # useful functions for dealing with PREDICTS data
library(roquefort) # useful PREDICTS functions, particularly for plotting
library(raster) # for dealing with spatial data
library(dplyr) # for handy data manipulation functions
library(tidyr) # ditto
library(lme4) # for mixed effects models
library(car) # for getting anova tables with significance values
library(DHARMa) # for model criticism plots
library(MuMIn) # for checking explanatory power of mixed effects models
library(stringr)
library(kableExtra)
```


*1. Explore number of sites*

```{r load data}

# Import site metrics of plants not dispersed by animals, plants dispersed by endozoochory and
# animals
diversity5 <- readRDS( file = "./output/cleaned_data/02_Statistical_Analysis_Site_metrics_combined_animals_endooPlants_notEndooPlants.rds")

```


Plants not dispersed by animals have enough sites to separate them into:

* Primary minimal, light and intense uses
* ISv minimal and light-intense uses
* Plantation forest minimal and light-intense uses
* Everything else with  merge all intensities


I will run in with those classes and then compare if it's better to merge everything.

```{r land use classes}

# Create a vector with the land uses categories that have enough sites to separate
# them into different intensities
LandUse_divide <- c("Primary", "ISV", "Plantation forest")

fourth_model_data <- diversity5 %>%
  # make a level of Primary minimal
  mutate(
    
    # collapse primary forest and non-forest together into primary vegetation as these 
    # aren't well distinguished
    Predominant_land_use = recode_factor(Predominant_land_use, 
                                         "Primary forest" = "Primary", 
                                         "Primary non-forest" = "Primary"),
    
    # indeterminate secondary veg and cannot decide are transformed into NA, urban too because 
    # it has only 40 sites. 
    Predominant_land_use = na_if(Predominant_land_use, "Secondary vegetation (indeterminate age)"),
    Predominant_land_use = na_if(Predominant_land_use, "Cannot decide"),
    Predominant_land_use = na_if(Predominant_land_use, "Urban"),
    
    
    # Give a shorter name to some land use categories 
    Predominant_land_use = str_replace_all(Predominant_land_use, 
                                           pattern = c("Young secondary vegetation" = "YSV",
                                                      "Intermediate secondary vegetation" = "ISV", 
                                                        "Mature secondary vegetation" = "MSV")),
    
    # Drop the Cannot decide intensity levels for the land-use categories that have 
    # enough sites for the minimal, and light/intense
    Use_intensity = ifelse(Predominant_land_use %in% LandUse_divide & Use_intensity == "Cannot decide",
                           NA,
                           paste(Use_intensity)), 
    
    # Join the intensity levels of light and intense for Plantation forest and ISV 
    Use_intensity = ifelse((Predominant_land_use == "Plantation forest" | 
                              Predominant_land_use == "ISV") &
                             (Use_intensity == "Intense use" | 
                                Use_intensity == "Light use"),
                           str_replace_all(Use_intensity, 
                                           pattern = c("Intense use" = "Light-intense use",
                                                                      "Light use" = "Light-intense use")),
                           paste(Use_intensity)),
    
    # Merge all the intensity levels for those land-use categories that don't have enough sites
    # in each land-use type/intensity combination
    Use_intensity = ifelse(Predominant_land_use %nin% LandUse_divide,
                           str_replace_all(Use_intensity, pattern = c("Intense use" = "All", 
                                                                      "Light use" = "All", 
                                                                      "Minimal use" = "All", 
                                                                      "Cannot decide" = "All")), 
                           paste(Use_intensity)),
    
    
    # Paste the land-use classes and intensity levels
    LandUse = ifelse(Predominant_land_use != "NA" & Use_intensity != "NA",
                     paste(Predominant_land_use, Use_intensity),
                     NA),
    
    # set reference level
    LandUse = factor(LandUse),
    LandUse = relevel(LandUse, ref = "Primary Minimal use")
    
  )


```

*2. Test for collinearity *

Since I'm going to explore the collinearity between categorical variables, I'm going to use 
the Generalized variance Inflation Factors function provided by Zuur et al., (2009)

```{r collinearity}

# Get the function
source("https://highstat.com/Books/Book2/HighstatLibV10.R")

# Calculate the VIF
corvif(fourth_model_data[ , c("LandUse", "Kingdom")])

```

*3. Complete cases*

```{r complete 1}

# Drop sites that don't have abundance measures or land-use data
fourth_model_data1 <- drop_na(fourth_model_data, 
                              Total_abundance, LandUse) %>% droplevels()


# Check the number of sites for animals and plants
addmargins(table(fourth_model_data1$LandUse, fourth_model_data1$Kingdom), 2)

```

*4. Transform rescaled abundance measures*

Abundance data usually display a nonnormal error distribution because they have a positive mean-variance relationship and are zero-inflated (Purvis et al., 2018).  Given that some abudance measures are not integers (some are relative abudance or densities), I am not going to model the abudance with a Poisson distribution, but I'm going to transform it in order to meet the assumptions of linear mixed models.

```{r transform 1}
# Transform RescaledAbundance.
fourth_model_data2 <- mutate(fourth_model_data1, 
                            logAbundance = log(RescaledAbundance + 1),
                            sqrtAbundance = sqrt(RescaledAbundance)
)


```

*5. Select random effects structure*


To select the random-effects structure, we use the method recommended by (Zuur et al., 2009) 
of taking the most complex fixed-effects structure, including all interactions, that will be 
tested in the second stage of modelling, and use it to compare different random-effects 
structures.

```{r m41}
m4_1 <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1|SS) + (1|SSB), data = fourth_model_data2)
```



```{r m42}
m4_2 <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1|SS) + (1|SSB) + (1|Source_ID), data = fourth_model_data2)
```


```{r m43, warning=TRUE}
m4_3 <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1+LandUse|SS) + (1|SSB), data = fourth_model_data2)
```


```{r m45, warning=TRUE}
m4_5 <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1+Predominant_land_use|SS) + (1|SSB), data = fourth_model_data2)
```



```{r AIC1}
# Compare the models that converged
AIC(m4_1, m4_2)
```


*6. Select fixed effect structure*

```{r anova 1}
# See the significance of the terms in the model.
Anova(m4_2)
```

The Land_use:kingdom interaction is significant, so I won’t remove this term from the model. 

*7. plot residuals*

```{r residuals 1}
simulationOutput4 <- simulateResiduals(fittedModel = m4_2, plot = TRUE)

```

*8. Model estimates*

Table with confidence intervals

```{r CI model 2}
CI_m4_2 <- read.csv("./output/summary_m4_2.csv", header = TRUE, sep = ";")
kable(CI_m4_2, format="latex", booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")
```

*9. Merge all land-use intensities*

```{r merge2}

# make a copy of the database: 
WH_merged <- fourth_model_data2

# Same model but with log 
m4_2l <- lmer(logAbundance~ LandUse + Kingdom + LandUse:Kingdom +
                (1|SS) + (1|SSB) + (1|Source_ID), data = WH_merged)

# Merge land-use intensities for the fourth_model_data2
WH_merged <- WH_merged %>% mutate(
  
  # Merge all the intensity levels for all land-use categories 
  Use_intensity = ifelse(Use_intensity != "NA", "All", "NA"),
  
  # Paste the land-use classes and intensity levels
  LandUse = ifelse(Predominant_land_use != "NA" & Use_intensity != "NA",
                   paste(Predominant_land_use, Use_intensity),
                   NA),
  
  # set reference level
  LandUse = factor(LandUse),
  LandUse = relevel(LandUse, ref = "Primary All")
  
  
)

# same model but with log
m4_2_ml <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
                  (1|SS) + (1|SSB) + (1|Source_ID), data = WH_merged)
```


Next, I will check if I’ve lost a significant amount of explanatory power by merging some land-use intensities.If I have, I want to keep the more complex model. If I haven’t lost a significant amount of explanatory power, then I can keep this simpler model. 

```{r}
anova(m4_2_ml, m4_2l, test = "F")
```


The LRT is significant, so by removing the interaction from the model, I did lose a significant amount of explanatory power. This means that I should keep the more complex model. 


*9. Plant contrast*

What I'm going to do is a Plant contrast, that means I'm firts going to model the complex model
(the one that assummes the reponse between endo and not-endo plants is significantly different), in this model I will  use the Kingdom interaction that has 3 levels: animals and endo plants and not endo plants. Then I am going to model the null model that assumes there is no difference between endo and not-endo plants. In this model the Kingdom will have 2 levels: animals, plants. Then I will compare both models using an ANOVA

```{r merge 3}
# make a copy of the database: 
WI_merged <- fourth_model_data2
```


```{r model 42 1}
# According to the last steps, our complex model is:
m4_2 <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1|SS) + (1|SSB) + (1|Source_ID), data = WI_merged)
```


```{r merge 4}
# For the null model I am going to replace the nePlantae for Plantae 
WI_merged <- WI_merged %>%
  
  # Replace nePlantae for Plantae
  mutate(
    Kingdom = recode_factor(Kingdom, "nePlantae" = "Plantae"), 

# set reference level
LandUse = factor(LandUse),
LandUse = relevel(LandUse, ref = "Primary Minimal use"), 
Kingdom = factor(Kingdom),
Kingdom = relevel(Kingdom, ref = "Animalia"))


```

```{r model s1}
m4_2_m <- lmer(logAbundance ~ LandUse + Kingdom + LandUse:Kingdom +
               (1|SS) + (1|SSB) + (1|Source_ID), data = WI_merged)
```

Compare complex and null model:

```{r model s2}
anova(m4_2_m, m4_2, test = "F")

```

*10. Plot the results of the selected abundance model*

```{r plot plot}
# Read r code from a file which contains the function to make the plot
source("./R/PlotErrBar_interactions.R")
source("./R/PlotErrBar_interactions_modified.R")

# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = m4_2, resp = "Abundance", Effect1 = "LandUse", Effect2 = "Kingdom",
                             ylims = c(-0.25,0.25), pointtype = c(16,17,18),blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), y = -0.18, labels = c("Primary minimal use",
                                            "Cropland all uses",
                                            "ISV light-intense use",
                                            "ISV minimal use",
                                            "MSV all uses",
                                            "Pasture all uses",
                                            "Plantation forest light-intense use",
                                            "Plantation forest minimal use",
                                            "Primary intense use",
                                            "Primary light use",
                                            "YSV all uses"), srt = 18, cex= 0.9)
```

