---
title: "Robust analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(error = TRUE)
library(knitr)
```

```{r, include=FALSE}
# Load libraries
library(influence.ME)
library(dplyr)
library(data.table)
library(broom)
library(lme4) 
library(robustlmm)

# Load the modified influence function
source("./R/04_Analise_Influence_modified_influenceME_function.R") 
# Add the Land use data
# Load function to merge LUI
source("./R/02_Statistical_Analysis_merge_LandUses_Intensities.R")
source("./R/PlotErrBar_interactions.R")
source("./R/PlotErrBar_interactions_modified.R")

```

```{r, include=FALSE}
# Vector of the land uses where intensity levels will remain separate
separate <- c("Primary", "Cropland", "ISV", "Plantation forest")

# Vector of land uses where light and intense uses will be merged
light_intense <- c("Primary", "Cropland", "ISV", "Plantation forest")
```

```{r, include=FALSE}

# Load abundance data
abundance <- readRDS("./output/cleaned_data/02_Statistical_Analysis_Abundance_Site_metrics_combined_animals_endooPlants_notEndooPlants.rds")

# Create the logAbundance column
abundance <- mutate(abundance, logAbundance = log(RescaledAbundance + 1))



# Merge LUI
abundance <- Merge_landUses_and_intensities(dataset = abundance, 
                                       index = 0, 
                                       land_uses_separate_intensities = separate, 
                                       land_uses_merge_light_intense = light_intense, 
                                       reference = "Primary Minimal use")


# Get complete cases
abundance <- drop_na(abundance, RescaledAbundance, LandUse.0) %>% droplevels()

```

**Total Abundance Model**

I used the influence.ME package to identify influential Source IDs. The first measured I assessed was the Cook's distance. The cut off for this measure is: 4/n, being n the number of groups, in this case 85. 
Cut off= 0.047

```{r, include=FALSE}
# Load results of influence.ME function
INFM1_abundance <- readRDS(file = "./output/intermediate_files/04_Analysis_influence_Abundance.Rds")
```


```{r, include=FALSE}
# Make a dataframe with the results of cooks distance
cooks_abundance <- as.data.frame(influence.ME::cooks.distance.estex(INFM1_abundance, sort = TRUE))
# Turn Source ID into a column
cooks_abundance <- setDT(cooks_abundance, keep.rownames = "Source_ID")
# Select the values of all columns that are greater for the cut off
cook_sources_abundance <- cooks_abundance %>% filter(V1 > 0.05) %>% droplevels

```


```{r }
# Count number of source IDs that are greater than the cut off
length(cook_sources_abundance$Source_ID)
```


```{r}
# Plot the results 
influence.ME::plot.estex(x=INFM1_abundance, which = "cook", cutoff = 0.047)
```

The second measure I assessed was the DFBETAS. The cut off for this measure is: 2/sqrt(n), being n the number of groups, in this case 85. 
Cut off= 0.217

```{r, include=FALSE}
# Make a dataframe with the results of dfbetas
dfbetas_abundance <- as.data.frame(influence.ME::dfbetas.estex(INFM1_abundance))
# Turn Source ID into a column
dfbetas_abundance <- setDT(dfbetas_abundance, keep.rownames = "Source_ID")
# Select the values of all columns that are greater for the cut off
dfbetas_sources_abundance <-dfbetas_abundance %>%  
  filter_if(is.numeric, any_vars(. > 0.22))
```


```{r }
# Count number of source IDs that are greater than the cut off for any parameter
length(dfbetas_sources_abundance$Source_ID)
```

**Species Richness model**

Cut off Cook's distance = 4/84 = 0.048


```{r, include=FALSE}

# Load species richness data
sp_richness <- readRDS("./output/cleaned_data/02_Statistical_Analysis_Richness_Site_metrics_animals_endooPlants_notendooPlants.rds")


# Merge LUI
sp_richness <- Merge_landUses_and_intensities(dataset = sp_richness, 
                                       index = 0, 
                                       land_uses_separate_intensities = separate, 
                                       land_uses_merge_light_intense = light_intense, 
                                       reference = "Primary Minimal use")


# Get complete cases
sp_richness <- drop_na(sp_richness, 
                         Species_richness, LandUse.0) %>% droplevels()

```


```{r, include=FALSE}
# Load results of influence.ME function
INFM1_richness <- readRDS(file = "./output/intermediate_files/04_Analysis_influence_Species_richness.Rds")

```


```{r, include=FALSE}

# Make a dataframe with the results of cooks distance
cook_richness <- as.data.frame(influence.ME::cooks.distance.estex(INFM1_richness, sort = TRUE))
# Turn Source ID into a column
cook_richness <- setDT(cook_richness, keep.rownames = "Source_ID")
# Select the values of all columns that are greater for the cut off
cook_sources_richness <- cook_richness %>% filter(V1 > 0.05) %>% droplevels

```


```{r}
# Count number of source IDs that are greater than the cut off
length(cook_sources_richness$Source_ID)
```


```{r E}
# Plot the results 
influence.ME::plot.estex(x=INFM1_richness, which = "cook", cutoff = 0.05)
```

Cut off DFBETAS = 2/sqrt(84) = 0.22

```{r, include=FALSE}
# Make a dataframe with the results of dfbetas
dfbetas_richness <- as.data.frame(influence.ME::dfbetas.estex(INFM1_richness))
# Turn Source ID into a column
dfbetas_richness <- setDT(dfbetas_richness, keep.rownames = "Source_ID")
# Select the values of all columns that are greater for the cut off
dfbetas_sources_richness <-dfbetas_richness %>%  
  filter_if(is.numeric, any_vars(. > 0.22))
```


```{r }
# Count number of source IDs that are greater than the cut off
length(dfbetas_sources_richness$Source_ID)
```


**Simpson's diversity**

Cut off Cook's distance = 4/73 = 0.055


```{r, include=FALSE}
# Load species richness data
simpson <- readRDS("./output/cleaned_data/02_Statistical_Analysis_Simpson_Site_metrics_animals_endooPlants_notendooPlants.rds")

# Merge LUI
simpson <- Merge_landUses_and_intensities(dataset = simpson, 
                                              index = 0, 
                                              land_uses_separate_intensities = separate, 
                                              land_uses_merge_light_intense = light_intense, 
                                              reference = "Primary Minimal use")


# Get complete cases
simpson <- drop_na(simpson, 
                       Simpson_diversity, LandUse.0) %>% droplevels()

# Transform Simpson's diversity index
simpson  <- simpson  %>%
  
  # Create a new column to transform the response variable with log
  mutate(log_one_over_D = log(Simpson_diversity))


```


```{r, include=FALSE}
# Load results of influence.ME function
INFM1_simpson <- readRDS(file = "./output/intermediate_files/04_Analysis_influence_Simpson.Rds")


```


```{r, include=FALSE}

cook_simpson <- as.data.frame(influence.ME::cooks.distance.estex(INFM1_simpson, sort = TRUE))
# Turn Source ID into a column
cook_simpson <- setDT(cook_simpson, keep.rownames = "Source_ID")
# Select the values of all columns that are greater for the cut off
cook_sources_simpson <- cook_simpson %>% filter(V1 > 0.06) %>% droplevels

```


```{r }
# Count number of source IDs that are greater than the cut off
length(cook_sources_simpson$Source_ID)
```


```{r }
# Plot the results 
influence.ME::plot.estex(x=INFM1_simpson, which = "cook", cutoff = 0.06)

```

Cut off DFBETAS = 2/sqrt(73) = 0.23

```{r, include=FALSE}
dfbetas_simpson <- as.data.frame(influence.ME::dfbetas.estex(INFM1_simpson))
# Turn Source ID into a column
dfbetas_simpson <- setDT(dfbetas_simpson, keep.rownames = "Source_ID")
# Select the values of all columns that are greater for the cut off
dfbetas_sources_simpson <-dfbetas_simpson %>%  
  filter_if(is.numeric, any_vars(. > 0.23))

```


```{r}
# Count number of source IDs that are greater than the cut off
length(dfbetas_sources_simpson$Source_ID)
```

*Robust analysis*

**Total abundance model**

```{r}
# Run the original model
full_model_abundance <- lmer(logAbundance ~ LandUse.0 + Kingdom + LandUse.0*Kingdom +
                               (1|Source_ID) + (1|SS) + (1|SSB), data= abundance)
```


```{r}

# Run the robust analysis
full_model_abundance_robust <- rlmer(logAbundance ~ LandUse.0 + Kingdom + LandUse.0*Kingdom +
                                       (1|Source_ID) + (1|SS) + (1|SSB), data = abundance)
```

Residuals of the robust model

```{r}
plot(full_model_abundance_robust)[1]
```


```{r}
plot(full_model_abundance_robust)[2]
```


```{r}
# Get the robustness weights for the sites
we_abundance <- getME(full_model_abundance_robust, "w_e")
# histogram of the robustness weights of sites
histogram(we_abundance)
```

Results of both models:

Original model

```{r,  echo=FALSE}
# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = full_model_abundance,
                             resp = "log(Rescaled Abundance)",
                             Effect1 = "LandUse.0", 
                             Effect2 = "Kingdom",
                             ylims = c(-0.25, 0.3),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -0.2, labels = c("Primary Minimal",
                          "Cropland Light-Intense",
                          "Cropland Minimal",
                          "ISV Light-Intense",
                          "ISV Minimal",
                          "MSV All",
                          "Pasture All",
                          "Plantation Light-Intense",
                          "Plantation Minimal",
                          "Primary Light-Intense",
                          "YSV All"),
     srt = 18, cex= 0.5)

```

Robust model

```{r,  echo=FALSE}
# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = full_model_abundance_robust,
                             resp = "log(Rescaled Abundance)",
                             Effect1 = "LandUse.0", 
                             Effect2 = "Kingdom",
                             ylims = c(-0.25, 0.3),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -0.2, labels = c("Primary Minimal",
                          "Cropland Light-Intense",
                          "Cropland Minimal",
                          "ISV Light-Intense",
                          "ISV Minimal",
                          "MSV All",
                          "Pasture All",
                          "Plantation Light-Intense",
                          "Plantation Minimal",
                          "Primary Light-Intense",
                          "YSV All"),
     srt = 18, cex= 0.5)

```


**Species Richness model**

```{r}
# Run the original model
full_model_richness <- glmer(Species_richness ~ LandUse.0 + Kingdom + LandUse.0:Kingdom +
                               (1|Source_ID) + (1|SS) + (1|SSB) + (1|SSBS), 
                             data = sp_richness, family = poisson, 
                             control = 
                               glmerControl(optimizer = "bobyqa", 
                                            optCtrl = list(maxfun = 20000)))
```


```{r}

# Run the robust analysis:
# Since the package only works for gaussian family, I'm going to model sp. richness as log(sp. richness + 1) 
# Create the logAbundance column
sp_richness <- mutate(sp_richness, logRichness = log(Species_richness + 1))

```


```{r}

full_model_richness_robust <- rlmer(logRichness ~ LandUse.0 + Kingdom + LandUse.0*Kingdom +
                                      (1|Source_ID) + (1|SS) + (1|SSB), data = sp_richness)
```

Residuals of the robust model

```{r}
plot(full_model_richness_robust)[1]
```


```{r}
plot(full_model_richness_robust)[2]
```


```{r}
# Get the robustness weights for the sites
we_richness <- getME(full_model_richness_robust, "w_e")
# histogram of the robustness weights of sites
histogram(we_richness)
```

Results of both models:

Original model

```{r,  echo=FALSE}
# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = full_model_richness,
                             resp = "log(Species Richness)",
                             Effect1 = "LandUse.0", 
                             Effect2 = "Kingdom",
                             ylims = c(-3, 2),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -3, labels = c("Primary Minimal",
                        "Cropland Light-Intense",
                        "Cropland Minimal",
                        "ISV Light-Intense",
                        "ISV Minimal",
                        "MSV All",
                        "Pasture All",
                        "Plantation Light-Intense",
                        "Plantation Minimal",
                        "Primary Light-Intense",
                        "YSV All"),
     srt = 18, cex= 0.5)
```

Robust model

```{r,  echo=FALSE}
# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = full_model_richness_robust,
                             resp = "log(Species Richness)",
                             Effect1 = "LandUse.0", 
                             Effect2 = "Kingdom",
                             ylims = c(-3, 2),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -3, labels = c("Primary Minimal",
                        "Cropland Light-Intense",
                        "Cropland Minimal",
                        "ISV Light-Intense",
                        "ISV Minimal",
                        "MSV All",
                        "Pasture All",
                        "Plantation Light-Intense",
                        "Plantation Minimal",
                        "Primary Light-Intense",
                        "YSV All"),
     srt = 18, cex= 0.5)

```

**Simpson's diversity model**

```{r}
# Run the original model
full_model_simpson <- lmer(log_one_over_D ~ LandUse.0 + Kingdom + LandUse.0:Kingdom +
                             (1|Source_ID) + (1|SS) + (1|SSB),
                           data = simpson)
```


```{r}

# Run the robust analysis:
full_model_simpson_robust <- rlmer(log_one_over_D ~ LandUse.0 + Kingdom + LandUse.0:Kingdom +
                                     (1|Source_ID) + (1|SS) + (1|SSB),
                                   data = simpson)

```


Residuals of the robust model

```{r}
plot(full_model_simpson_robust)[1]
```


```{r}
plot(full_model_simpson_robust)[2]
```


```{r}
# Get the robustness weights for the sites
we_simpson <- getME(full_model_simpson_robust, "w_e")
# histogram of the robustness weights of sites
histogram(we_simpson)
```

Results of both models:

Original model

```{r,  echo=FALSE}
# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = full_model_simpson,
                             resp = "log(Simpson's diversity index)",
                             Effect1 = "LandUse.0", 
                             Effect2 = "Kingdom",
                             ylims = c(-1, 1),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -1, labels = c("Primary Minimal",
                        "Cropland Light-Intense",
                        "Cropland Minimal",
                        "ISV Light-Intense",
                        "ISV Minimal",
                        "MSV All",
                        "Pasture All",
                        "Plantation Light-Intense",
                        "Plantation Minimal",
                        "Primary Light-Intense",
                        "YSV All"),
     srt = 18, cex= 0.5)

```

Robust model

```{r,  echo=FALSE}
# Plot the differences between estimates 
PlotErrBar_interactions_modi(model = full_model_simpson_robust,
                             resp = "log(Simpson's diversity index)",
                             Effect1 = "LandUse.0", 
                             Effect2 = "Kingdom",
                             ylims = c(-1, 1),
                             pointtype = c(16,17, 18),
                             blackwhite = FALSE)

# Plot the x label
text(x = c(0.8:10.8), 
     y = -1, labels = c("Primary Minimal",
                        "Cropland Light-Intense",
                        "Cropland Minimal",
                        "ISV Light-Intense",
                        "ISV Minimal",
                        "MSV All",
                        "Pasture All",
                        "Plantation Light-Intense",
                        "Plantation Minimal",
                        "Primary Light-Intense",
                        "YSV All"),
     srt = 18, cex= 0.5)


```